---
title: Introducing flxr - an R package to retrieve movie data
author: Mark Roepke
date: '2018-02-23'
slug: introducing-flxr-an-r-package-to-retrieve-movie-data
comments: false
categories:
  - r
  - movies
  - scraping
  - packages
tags:
  - r
  - movies
  - scraping
  - packages
---

I am happy to say that I've created my first R package: `flxr`. This package is designed to scrape basic data about movies.

It can be installed using the `devtools` package:

```{r eval = FALSE}
devtools::install_github("markroepke/flxr")
library(flxr)
```

```{r echo = FALSE}
library(flxr)
```

The `flxr` package retrieves information about movies using only a user-provided movie title. For example, if we wanted to find the director of _The Godfather_, we could use the `flx_mov_director()` function:

```{r}
flx_mov_director(movie = "The Godfather")
```

You will notice the message above specifying that the director for _The Godfather_ is being returned. This is printed to ensure the user that the correct movie was found based on the movie name provided to the function.

There are many other functions available in `flxr` to retrieve other data. These functions include:

* `flx_mov_desc()` to return the movie description
* `flx_mov_genre()` to return the movie genre
* `flx_mov_runtime()` to return the movie runtime
* `flx_mov_user_rtg()` to return the movie average IMDB user rating
* `flx_mov_mc_rtg()` to return the movie Metacritic rating
* `flx_mov_lead_cast()` to return the movie title cast

<br>

And finally, there is a function to return all of the above information:

```{r}
library(magrittr)
flx_mov_details(movie = "The Godfather") %>%
  knitr::kable()
```

<br>

I am hoping to use this package in the coming days to look into the Oscar-nominated films.